Efficient Parallelization of $\ell_1$-Minimization for Face Recognition \thanks{Corresponding author: . This work was partially supported by ARO MURI W911NF-06-1-0076.}

### 2010-10-21 paper/main.tex
Pulled in my notes on results from a synthetic benchmark we did.

>>>>

\section{Benchmarks on Synthetic Data}
\subsection{PALM L1 minimization benchmark}

Optimization:
\begin{equation}
\min_{\x, \e} \|\e\|_1 \quad \subj \quad \y = A \x + \e.
\end{equation}

$A$ is $mxn$
$m = 5120$
$n = 32$

One of these minimizations is performed per 
alignment iteration in our face recognition pipeline.

All of the following are implementations of primal ALM.
Primal ALM contains outer and inner iterations.
To simplify the benchmark, the number of outer iterations
is clamped to 50, and the number of inner iterations is
clamped to 50 iterations per outer loop.  Running with
a convergence criterion will typically result in fewer
iterations, depending on the data, so these numbers are
for comparison across hardware, and may look worse than
with other benchmarks on face data.

To generate a meaningful comparison between different hardware
architectures, we compare on a per-board basis:

For CPU implementations, the benchmark makes use of
as all of the cores in as many cpu's are present.

For GPU implementations, the benchmark makes the best use
of the entire GPU chip (most GPU boards have a single GPU chip).

Since different platforms may perform most efficiently when
executing multiple L1 minimizations simultaneously,
we quote the average time required to solve a single L1
minimization problem instance.

schur:  Mac Pro, dual socket, quad core, hyperthreading intel nehalem @ 2.66 GHz

Nvidia GTX480: This is the current state of the art consumer card.  
It is clocked slightly faster than the pro version of the card, so it 
should have slightly faster single precision float performance. (double precision is crippled, though)
GTX480 has 15 cores, each of which has 2x32-wide floating point pipelines.
All Fermi GPU's have 768KB L2 cache.

Zihan's OpenCV implementation: *** 190 ms per L1 minimization ***
amortized over 16 instances,
four of which run simultaneously.  Tuning note:  the cpus
can support up to 16 logical threads, but 4 instances seem
sufficient to saturate the memory architecture.  Using more
than 5 instances results in lower performance, perhaps due
to less efficient usage of the cpu caches.

Zihan's MATLAB implementation: *** 60 ms per L1 minimization ***
This is amortized over 100 sequential instances.  All 16 logical
cores were used, i.e. showed 100% usage in htop.  Pararllelism
is a result of MATLAB's usage of Intel's MKL, a parallel blas
implementation. I believe performance is very high to
efficient cache usage:

The A matrix takes up 5120*32*8 bytes = 1.25 MB
The two processors have 16MB L3 Cache in total.
The two processors have 2MB L2 Cache in total.

Drew's pure CUDA, single kernel implementation: *** 36 ms per L1 minimization ***
This implementation performs L1 minimization on 60 instances of L1.
They are all run as a single cuda kernel.  
Each instance of L1 is mapped onto a cuda block. (tuned to 64 threads / block)
This means that the GPU hardware will schedule the multiple instances to run
on the (15 for GTX480) cuda cores as concurrently (15? 30?) as possible.  
For this to work, all of the data has to be ready at the same time.

Drew's pure CUDA, streams implementation: *** 70 ms per L1 minimization ***
This implementation performs L1 minimization on 60 instances of L1.
Each instance of L1 is mapped onto a cuda block. (tuned to 128 threads / block)
Each instance is run in a separate single-block CUDA kernel.
The 60 kernels are added to 15 cuda streams.
The GPU (hardware? driver?) then runs the kernels, up to 15 concurrently.

Victor's CUDA/cublas mixed implementation: *** 302 ms per L1 minimization ***
This implementation makes use of hand-written kernels for vector-vector operations,
and cublas for matrix-vector operations.  There are two implementations under development:
1. Perform a single L1 minimization using the entire GPU.
2. Use cuda streams to perform multiple L1 minimizations concurrently on the GPU.
Benchmarked on Drew's GTX480

Drew's OpenMP implementation: *** 200 ms per L1 minimization***
Benchmarked on schur(Dual Socket Quad Core Nehalem), all compiler optimizations turned on.
This divides the matrix operations into 8 chunks, each of which is handled by a thread
on a different CPU core.  Fine-grained parallelism is achieved through SSE2 vector
floating point operations.  This implementation should ultimately beat the MATLAB/MKL
implementation, especially for small problem sizes, since the primary advantage
over MKL is a reduction in the number of times image-sized temporary variables
are written and read from memory.

Drew's Single-Threaded MKL implementation: *** 226 ms per L1 minimization ***
Benchmarked on schur(Dual Socket Quad Core Nehalem), all compiler optimizations turned on.
Uses MKL for single-threaded matrix-vector multiplies.  
Saves ~almost as many writes of temporary vector variables as the previous OpenMP implementation.
Shrinkage operator improved; at least with g++ there is now no branching or function call overhead.
This implementation uses a single CPU core...
... so if linear speedup can be achieved, could possibly get *** 28 ms per minimization ***
